---
title: "Introduction"
author: "auth"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=F}
##### ONLY FOR TESTING ####
library(R6)
fl <- c("../R/filter.R" ,"../R/scBMClass.R")
for (f in fl) { source(f) }
###

require(Seurat)
dt1 <- Read10X(data.dir = "../../data/pbmc/filtered_gene_bc_matrices/hg19/")
so <- CreateSeuratObject(counts=dt1, project="test", min.features=200, min.cells=5)
```

# Outline
```{r}
fl <- c("../R/filter.R" ,"../R/scBMClass.R")
for (f in fl) { source(f) }

####
# Create scBM instance w/ Seurat object so
obj <- scBM$new(so)

####
# Apply default filters 
obj <- filter(obj)
#obj <- filter(obj, ...)

#######
# Identify clusters 
# ... 
# ... 

######
# Extract cluster metrics  -> aggregated metrics for evaluation 
# .. # clusters, cluster size, ...

######
#####
# Assign identity to clusters




```

## Pseudocode f. cutoff detection

Init:
Filter_Number_of_ genes_filter =xx;  // any threshold , e.g. from literatur

Derive Threshold:
Computer_Mean_Nummber_of_genes (From profiles of all cells in all samples);
Computer_SDev_Nummber_of_genes (From profiles of all celles in all samples);

// Instead of mean (gauss distrbuted) :
// derive median instead of mean + mad or Interquartile distance instead of sdev

Filter_Number_of_genes = Compute_Mean_Nummber_of_genes + Compute_SDev_Nummber_of_genes // ~5% filter

//In case of very divergent data: Sdev+Mean > Total number  of genes
if Filter_Number_of_genes>Total_Numer of genes => Filter_Number_of_genes = Total_Numer of genes*0.95 

Filter:
Loop over all profiles
         Remove_Profiles  where Number_of_ genes < Filter_Number_of_genes

To detect outlier samples:
Compute grandmean/grandsdev for all samlpes/cells  as well as Sample_i_Mean/Sampe_i_Sdev for each sample on its specidv cell profiles.

Find sample where p-tests grandmens from t-test(Granmean/grandsdev vs. Samplev_Mean/DSdev) <0.05 
Falg or Exclude such sample or handle indivdually




##########

# Overview
This vignette demonstrates the use of the scBM package.

# Requirements and installation
The package utilizes Seurat. 


