---
title: "ScRNA_seq"
output: html_document
date: "2023-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ScRNA seq analysis.

This document gives an overview of the scRNA analysis using Seurat for clustering and expression of genes of interest.
This file covers an overview of Preprocessing steps, Normalization, Dimensionality reduction, clustering, and expression of target genes

Loading packages
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(readr)
library(tidyverse)
library(plotly)
```


## 1. Preprocessing
After creation of the Seurat object, Cells with low feature and count number and high levels of mitochondrial RNA are eliminated. 
```{r}
t1 <- readRDS("/filename.rds")

#Preprocessing and visualization 
t1[["percent.mt"]] <- PercentageFeatureSet(t1, pattern = "^MT-")
t1[["log10Genes.per.UMI"]] <- log10(t1$nFeature_RNA) / log10(t1$nCount_RNA)
VlnPlot(t1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

t1 <- c(t1)
preprocessing = function(x) {
  median_ncount <- median(x$nCount_RNA)
  median_nfeatures <- median(x$nFeature_RNA)
  median_mt <- median(x$percent.mt)
  
  mad_ncounts <- mad(x$nCount_RNA, center = median_ncount)
  mad_nfeatures <- mad(x$nFeature_RNA, center = median_nfeatures)
  mad_mean_mt <- mad(x$percent.mt, center = median_mt)
  
  total_ncounts <- median_ncount+mad_ncounts
  total_nfeatures <- median_nfeatures+mad_nfeatures
  total_mt <- median_mt+mad_mean_mt
  
  x@meta.data %>% 
    ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") + 
    scale_x_log10() + 
    scale_y_log10() + 
    theme_classic() +
    geom_vline(xintercept = total_ncounts) +
    geom_hline(yintercept = total_nfeatures) +
    geom_vline(xintercept = 500) +
    geom_hline(yintercept = 250) +
    facet_wrap(~orig.ident)
  
  x<- subset(x, subset = (nFeature_RNA >= 250) & 
                  (nFeature_RNA <= total_nfeatures) &
                  (nCount_RNA >= 500) & 
                  (nCount_RNA <= total_ncounts) &
                  (percent.mt < total_mt) &
                  (log10Genes.per.UMI > 0.8))
 
}

preprocessed.obj <- lapply(t1, preprocessing)
preprocessed.obj
```

## 2. Normalization
Preprocessed data can now be Normalized and scaled for compensating the variance of highly variable genes. 
Methods for Normalization include 'LogNormalize', 'CLR' (Centered log ratio), 'RC' (Relative Counts), which can be employed by changing the normalization.method. Default is LogNormalize
Finding Variable features also include different selection methods 'vst' (local polynomial regression), 'mvp' (mean.var.plot), and 'disp' (dispersion). Number of variable features is automatically selected for mvp
Expression of Housekeeping genes is checked for viable cells

```{r}
#Normalization and Feature Selection
GOI <- c("TGFB1", "TGFB2", "TGFB3", "TGFBR1", "TGFBR2", "TGFBR3", "SMAD1", "SMAD2", "SMAD3", "SMAD4", "SMAD7", 
         "PIK3CA", "TNF", "TRAF6", "TRAF4", "NFKB1", "MAP3K7", "MAPK14", "MAPK1", "MAPK3", "MAPK8")
#If only ensembl ID is available
GOI_Ensembl <- c("ENSG00000228978", "ENSG00000223952", "ENSG00000175387",  "ENSG00000100030", "ENSG00000119699", "ENSG00000101665", " ENSG00000141646",  "ENSG00000069702", "ENSG00000163513", "ENSG00000102882", "ENSG00000092969", "ENSG00000107643", "ENSG00000105329", "ENSG00000170365", "ENSG00000166949", "ENSG00000109320", "ENSG00000106799", "ENSG00000121879", "ENSG00000232810", "ENSG00000076604", "ENSG00000135341", "ENSG00000112062", "ENSG00000175104")

#Normalization 
normalized.obj <- NormalizeData(preprocessed.obj , normalization.method = "LogNormalize", scale.factor = 10000)
normalized.obj  <- FindVariableFeatures(normalized.obj, selection.method = "vst", nfeatures = 2000)
all_vg_1 <- VariableFeatures(normalized.obj)
all_genes <- rownames(normalized.obj)
top_10 <- head(VariableFeatures(normalized.obj), 10)
top_10
capture.output(top_10, file = "Top_10_genes.txt")

#Visualization of highly variable genes
plot1 <- VariableFeaturePlot(normalized.obj)
plot2 <- LabelPoints(plot = plot1, points = top_10, repel = TRUE)
plot1 + plot2

common <- intersect(all_vg_1, GOI_Ensembl)
common

plot3 <- LabelPoints(plot = plot1, points = common, repel = TRUE)
plot1 + plot3

#Scaling data
normalized.obj <- ScaleData(normalized.obj, features = all_genes)
normalized.obj
head(normalized.obj@meta.data,5)

#Checking for HK genes expression
HK <- FetchData(normalized.obj, vars = c("GAPDH", "ACTB"))
expr <- colSums(HK)
HK_expr <- rowSums(HK)
which(HK_expr == 0)
```

## 3. Dimensionality reduction
The method of choice can either be linear (PCA) or non-linear (UMAP or tSNE) by changing the respective parameters. UMAP can either be generated for selected features or using dimensions of PCA (features set to NULL and use dims = )

```{r}
reduction.obj <- RunPCA(normalized.obj, npcs = 100)
reduction.obj@reductions$pca@cell.embeddings
reduction.obj <- RunUMAP(reduction.obj, umap.method = "umap-learn", reduction.name = "umap", n.components = 4, metric = "correlation", features = VariableFeatures(reduction.obj), verbose = TRUE, min.dist = 0.9)
DimPlot(reduction.obj, reduction = "umap", dims = c(1,2))
```

For selecting number of dimensions one can use jackstraw and/or elbowplot. It must be noted that jackstraw takes more processing time for large datasets.

```{r}
reduction.obj <- JackStraw(reduction.obj, num.replicate = 100)
reduction.obj <- ScoreJackStraw(reduction.obj, dims = 1:20)
JackStrawPlot(reduction.obj, dims = 1:15)

ElbowPlot(reduction.obj)
```

## 4. Clustering
Neighbouring cells can be found either using number of dimensions or variable features from PCA or UMAP.
Clusters can be found using louvain or leiden algorithm and by varying the resolution (0.4 to 1.4 for large data would be ideal)

3D visualization is possible with plotly function

```{r}
reduction.obj<- FindNeighbors(reduction.obj, features = VariableFeatures(reduction.obj), reduction = "umap", dims = 1:4)
clustered.obj <- FindClusters(reduction.obj, resolution = 0.5, algorithm = "louvain")
head(Idents(clustered.obj), 5)
UMAPPlot(clustered.obj)

#3d visualization
Embeddings(object = clustered.obj, reduction = "umap")
plot.data <- FetchData(object = clustered.obj, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "seurat_clusters"))
plot.data$label <- paste(rownames(plot.data))
fig <- plot_ly(data = plot.data, 
               x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
               color = ~seurat_clusters
               type = "scatter3d",
               mode = "markers",
               marker = list(size = 5, width=2))
               
fig
```

## 5. Cluster annotation and target gene expression
Cells can be manually annotated using literature and available datasets and the clusters can be re-labeled. new.cluster.ids is used to save the respective cell annotation and then changed in the object

```{r}
cell.markers <- FindAllMarkers(clustered.obj, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
cluster_markers <- as.data.frame(cell.markers %>%
  group_by(cluster) %>%
  slice_max(n = 10, order_by = avg_log2FC))

new.cluster.ids <- c("Naive CD4 T", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
    "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(clustered.obj)
pbmc <- RenameIdents(clustered.obj, new.cluster.ids)

#Visualization
FeaturePlot(clustered.obj, features = GOI)


```

Target gene expression can also be visualized using cellscatter(), violin plots, etc. 
